---
title: "ASSIGNMENT 1"
author: "Krishna Krupa"
date: "2023-09-09"
output:
  html_document: default
  word_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

```

#Downloading a Dataset 

 https://www.kaggle.com/datasets/adityaab1407/employee-productivity-and-satisfaction-hr-data

#Importing the Dataset into R

```{r}
employee = read.csv("hr_dashboard_data.csv", header = TRUE, sep = ",") 
head(employee)

```

#Descriptive statistics of Quantitative variable

Mean, Median, Min, Max and Standard deviation of Dataset

```{r}
summary(employee$Age)
summary(employee$Projects_Completed)
summary(employee$Productivity)
summary(employee$Satisfaction_Rate_Per)
summary(employee$Feedback_Score)
```

#Descriptive statistics of categorical variable

1.Number of male and female in the Dataset

```{r}
 table(employee$Gender)
```
 
2.Proportion of each department in the Dataset

 ```{r}
count=table(employee$Department)
 proportion= prop.table(count)
 print(proportion)
 ```

#Transformation of Quantitative Variable

converting the feedback values into percentage

```{r}
result= (employee$Feedback_Score/5)*100
print(result)
```

#Plotting the variables

1.Bar plot for number of departments in the Dataset

```{r}
x=table(employee$Department)
print(x)

barplot.default(x,names.arg = c('Finance','HR','IT', 'Marketing', 'Sales' ),
                xlab= 'Departments', ylab = 'Number', ylim = c(0,50),main = 'Bar Graph',col = 'blue' )

```

2.Pie Chart for the number of positions in the data set

```{r}
x=table(employee$Position)

 print(x)
 
 pie(x, labels= names(x),radius = 1, clockwise = TRUE, main = 'Pie Chart' )
```

#Scatter plot for age and the number of projects completed in the Dataset 

```{r}
Age= employee$Age
projects= employee$Projects_Completed
plot(x = Age, y = projects, xlab='Age', ylab= 'No.of projects', xlim = c(0,100), ylim = c(0,50), main = 'scatterplot')